---
title: "Assign. 1 STA 445"
author: "Paige Hawkinson"
date: "`r Sys.Date()`"
output: pdf_document
---

## Directions:
This assignment covers chapter 5.  Please show all work in this document and knit your final draft into a pdf.  This is assignment is about statistical models, which will be helpful if you plan on taking STA 570, STA 371, or STA 571. 

## \color{black} Problem 1: Two Sample t-test

\color{blue} 

a. Load the `iris` dataset. 
```{r}
data("iris")
head(iris)
```


b. Create a subset of the data that just contains rows  for the two species setosa and versicolor using filter.  Use slice_sample to print out 20 random rows of the dataset.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
iris.subset <- iris %>%
  filter(Species == "setosa" | Species == "versicolor")
slice_sample(iris.subset)
```


c. Create a box plot of the petal lengths for these two species using ggplot. Does it look like the mean petal length varies by species?
```{r}
library(ggplot2)
ggplot(data=iris.subset, aes(x=Species, y=Petal.Length)) +
  geom_boxplot() +
  labs( title='Species vs Petal Length') +
  labs( x="Species", y="Petal Length")
```
Based on the box plot, it looks like there is a difference in mean petal length 

d. Do a two sample t-test using t.test to determine formally if the petal lengths differ. Note:  The book uses the tidy function in the broom package to make the output "nice".  I hate it!  Please don't use tidy.
```{r}
t.test(data=iris.subset, Petal.Length ~ Species)
```

d. What is the p-value for the test? What do you conclude?

The p-value for the t-test is 2.2e-16, or essentially zero. The p-value means we can reject the null hypothesis and that there is enough evidence to show that there is a difference in mean petal lengths between setosa and versicolor.
 
e. Give a 95% confidence interval for the difference in the mean petal lengths. 

We are 95% confident that the difference in mean petal lengths between setosa and versicolor is between -2.939618 cm and -2.656382 cm.  

f. Give a 99% confidence interval for the difference in mean petal lengths.  (Hint: type ?t.test.  See that you can change the confidence level using the option conf.level)
```{r}
t.test(data=iris.subset, Petal.Length ~ Species, conf.level=0.99)
```
We are 99% confident that the difference of mean petal lengths between setosa and versicolor is between -2.986265 cm and -2.609735 cm. 

g. What is the mean petal length for setosa?

The mean petal length for setosa is 1.462 cm.
 
h. What is the mean petal length for versicolor?

The mean petal length for versicolor is 4.260 cm.
 

## Problem 2: ANOVA 

Use the iris data with all three species.

a. Create a box plot of the petal lengths for all three species using ggplot.Does it look like there are differences in the mean petal lengths?
```{r}
library(ggplot2)
ggplot(data=iris, aes(x=Species, y=Petal.Length)) + 
  geom_boxplot() + 
  labs( title='Species vs Petal Length') +
  labs( x="Species", y="Petal Length")
```
Based on the box plot form above, there seems to be a difference in mean petal lengths among the three different species. 

b. Create a linear model where sepal length is modeled by species. Give it
an appropriate name.
```{r}
iris.mod.1 <- lm(data=iris, Sepal.Length ~ Species-1)
```


c. Type anova(your model name) in a code chunk.
```{r}
anova(iris.mod.1)
```

d. What is the p-value for the test?  What do you conclude.

The p-value is 2.2e-16 or essentially zero. This p-value means that we can reject the null hypothesis and that there is enough evidence to show that there is a difference in mean sepal length between species.  

e. Type summary(your model name) in a code chunk.
```{r}
summary(iris.mod.1)
```


f. What is the mean sepal length for the species setosa?

The mean sepal length for setosa is 5.0060 cm.

g. What is the mean sepal length for the species versicolor?

The mean sepal length for versicolor is 5.9360 cm. 


## Problem 3: Regression 

Can we describe the relationship between petal length and petal width?

a. Create a scatterplot with petal length on the y-axis and petal width on the x-axis using ggplot.
```{r}
ggplot(data=iris, aes(x=Petal.Width, y=Petal.Length, color=Species)) +
  geom_point() +
  labs( title='Petal Width vs Petal Length') +
  labs( x="Petal Width", y="Petal Length")
```


b. Create a linear model to model petal length with petal width (length is the response variable and width is the explanatory variable) using lm.
```{r}
iris.mod.2 <- lm(data=iris, Petal.Length ~ Petal.Width)
iris.mod.2
```


c. What is the estimate of the slope parameter?

Estimate of slope parameter = 2.230

d. What is the estimate of the intercept parameter?

Estimate of intercept parameter = 1.084

e. Use summary() to get additional information.
```{r}
summary(iris.mod.2)
```

## Problem 4: Modeling Trees

Using the `trees` data frame that comes pre-installed in R, follow the steps below to fit the regression model that uses the tree `Height` to explain the `Volume` of wood harvested from the tree.

a. Create a scatterplot of the data using ggplot.
```{r}
ggplot(data=trees, aes(x=Height, y=Volume, color=Girth)) +
  geom_point() +
  labs( title='Tree Height vs Tree Volume') +
  labs( x="Height (ft)", y="Volume (cubic ft)")
```


b. Fit a `lm` model using the command `model <- lm(Volume ~ Height, data=trees)`.
```{r}
model <- lm(Volume ~ Height, data=trees)
```


c. Print out the table of coefficients with estimate names, estimated value, standard error, and upper and lower 95% confidence intervals.
```{r}
summary(model)
confint(model)
```


d.Add the model fitted values to the `trees` data frame along with the regression model confidence intervals.  Note: the book does this in a super convoluted way.  Don't follow the model in the book.  Instead try cbind.
```{r}
trees.w.pred <- cbind(trees, predict(model, interval="confidence"))
trees.w.pred
```


e. Graph the data and fitted regression line and uncertainty ribbon.
```{r}
ggplot(data = trees.w.pred, aes(x=Height, y=Volume)) +
  geom_point() +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "lightgreen") +
  labs( title='Tree Height vs Tree Volume') +
  labs( x="Height (ft)", y="Volume (cubic ft)")
```


f. Add the R-squared value as an annotation to the graph using annotate.
```{r}
ggplot(data = trees.w.pred, aes(x=Height, y=Volume)) +
  geom_point() +
  geom_line(aes(y=fit)) +
  geom_ribbon(aes(ymin=lwr, ymax=upr), alpha=0.5, fill = "lightgreen") +
  geom_text(data = trees.w.pred, aes(x=82, y=72, label="R^2=0.3579")) +
  labs( title='Tree Height vs Tree Volume') +
  labs( x="Height (ft)", y="Volume (cubic ft)") 
```



