---
title: "STA_445_Assignment_6"
author: "Paige Hawkinson"
date: "2024-03-28"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(mosaicData)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.
```{r}
mdy('September 13, 2010')
```
b. Sept 13, 2010.
```{r, Eval=FALSE}
mdy('Sept 13, 2010')
```
We cannot run the code as R does not recognize Sept to be September. We would need to rewrite this to either Sep 13, 2020 or September 13, 2010.  

c. Sep 13, 2010.
```{r}
mdy('Sep 13, 2010')
```

d. S 13, 2010. Comment on the month abbreviation needs.
```{r, Eval=FALSE}
mdy('S 13, 2010')
```
This is similar to 1b where R does not recognize S as September. We would need to rewrite this in order for R to create the output we want. 

e. 07-Dec-1941.
```{r}
dmy('07-Dec-1941')
```
f. 1-5-1998. Comment on why you might be wrong.
```{r}
mdy('1-5-1998')
```
I might be wrong as I understood the date to be January 5th, 1998 but the date could also be day, month,  year... May 1st, 1998. 

g. 21-5-1998. Comment on why you know you are correct.
```{r}
dmy('21-5-1998')
```
I know I am correct because there are not 21 months, meaning the order is day, month, year. 

h. 2020-May-5 10:30 am
```{r}
ymd_hm('2020-May-5 10:30 am')
```


i. 2020-May-5 10:30 am PDT (ex Seattle)
```{r}
ymd_hm('2020-May-5 10:30 am', tz='US/Pacific')
```


j. 2020-May-5 10:30 am AST (ex Puerto Rico)
```{r}
ymd_hm('2020-May-5 10:30 am', tz='America/Puerto_Rico')
```

### Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:

a. Calculate the date of your 64th birthday.
```{r}
dob <- mdy('July 29, 2001')
dob + years(64)
```
b. Calculate your current age (in years). 
```{r}
todays.date <- Sys.Date()
current.age <- interval(dob, todays.date)
(current.age.2 <- year(as.period(current.age)))
```

c. Using your result in part (b), calculate the date of your next birthday.
```{r}
(next.bday <- years(current.age.2 +1) + dob)
```


d. The number of _days_ until your next birthday.
```{r}
as.period(todays.date %--% next.bday, unit = "days")
```


e. The number of _months_ and _days_ until your next birthday.
```{r}
as.period(todays.date %--% next.bday, unit = "months")
```

## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 
```{r}
AZ.time <- ymd_hm('2015-May-8 03:00 pm', tz='US/Arizona')
with_tz(AZ.time, "Pacific/Auckland")
```

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 
```{r}
new.Births78 <- Births78 %>%
  select("date", "births")
```


b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?
```{r}
ggplot(data = new.Births78, aes(x = date, y = births)) +
  geom_point() +
  labs( title = 'Number of births vs date in 1978') +
  labs( x = "Date", y = "Number of Births in integer")
```
In the graph, there is a clear distinction in the data points, causing there two be two groups of samples. This trend could be due to the day of the week the baby was born on. 

c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.
```{r}
new.new.Births78 <- new.Births78 %>%
  mutate(dow = wday(date, label = TRUE, abbr = FALSE))
```


d. Plot the data with the point color being determined by the day of the week variable.
```{r}
ggplot(data = new.new.Births78, aes(x = date, y = births, color=dow)) +
  geom_point() +
  labs( title = 'Number of births vs date in 1978') +
  labs( x = "Date", y = "Number of Births in integer")
```










